---
title: "Wk5_CorrectMultipleTests"
author: "Isaac Sowah Badu"
date: "2024-09-25"
output: pdf_document
---

```{r,include = TRUE, echo= TRUE}
library(knitr)
opts_knit$set(root.dir = "C:/Users/sowai23/Desktop/Year_2_First sem/R stats/Files/R-class-2024/WK05_Correctmultiple_Tests")
```



```{r}
# Defining a function
my.fun <- function(x){
y <- x*5 + 1
return(y)
}


# Using the function created
my.fun(10)
```


```{r}
# A function with two parameters
my.fun2 <- function(x, a){
y <- x*5 + 3^a
return(y)
}
my.fun2(10, 5)

```


```{r}
# What happens when you switch them around with, and without, specification? I got the same answers with or without specification
my.fun2(a=5, x=10)
## [1] 293
```


```{r}
my.fun3 <- function(x){
y <- x*5 + 3^a
return(y)
}
```

#Checkpoint 1: Call my.fun2(10) and my.fun3(10): what happens? Now, specify a <- 5, and call my.fun2(10) and myfun3(10) again. Which one works and why? What if x is also an object in R?


```{r}
##my.fun2(10)
```


```{r}
#my.fun3(10)
```

#my.fun2(10) says argument "a" is missing, with no default and my.fun3(10) says object 'a' not found

```{r}
##specifying 
a <- 5
```



```{r}
my.fun2(10, 5)
```


```{r}
##my.fun3(10, 5)
```


#my.fun2(10) now works but my.fun3(10) does not work. my.fun2(10) works now because we have given it an argument 5 which was initially missing. my.fun3(10) does not work because the function does not have an object a.


```{r}
my.fun3 <- function(x){
y <- x*5 + 3^a
return(y)
}
```


```{r}
#Writing the function called Bonf()
Bonf <- function(p.vect) {
  bonf1 <- 0.05 / length(p.vect)
  return(data.frame(p.vect, Is.Sig= p.vect < bonf1))
}

#usage
p.vect <- c(0.23, 0.02, 0.013, 0.045)
Bonf(p.vect)
```


```{r}
#Benjamini-Hotchberg
BenHotch <- function(p.vect){
rank.p <- rank(p.vect)
n <- length(p.vect)
rank.alpha<-rank.p*0.05/n
sig.p <- max(p.vect[p.vect<rank.alpha])
return(data.frame(p.vect, Is.Sig = p.vect<=sig.p))
}
```


```{r}
p.vect <- c(0.23, 0.02, 0.013, 0.045) 
BenHotch(p.vect)
```


```{r}
#if (!require("BiocManager", quietly = TRUE))
#install.packages("BiocManager")
#BiocManager::install("qvalue")
```
```{r}
library(qvalue)
```



```{r,include = TRUE, echo= TRUE}
library(knitr)
opts_knit$set(root.dir = "C:/Users/sowai23/Desktop/Year_2_First sem/R stats/Files/R-Class-2024/WK05_Correctmultiple_Tests")
```


```{r}
###Uploading the data
df.p <- read.table("data/1R_P30_1351142954_453_1_NumPops=30_NumInd=20Bayenv2LFMMpca-1.Cpval", header=TRUE)
```



```{r}
# A logical indicating if a locus is neutral (TRUE)
IsNeut <- df.p$IsNeut[df.p$UseSNP] 

# The p-value for each locus calculated by a program
p.list <- df.p$K5.P[df.p$UseSNP] 

### We can check the assumption that our p-values are uniformly distributed
### by plotting a histogram:
hist(df.p$K5.P)
```



```{r}
##Comparing figure 1 in Storey and Tibshirani (2003), where there is only a small inflation at p<0.01.
### Compare it to a histogram of p-values from a normal distribution:
hist(pnorm(rnorm(1000)))
```



```{r}
### Making a logical list of whether the q-values are <0.05
q.out <- qvalue(p.list)$qvalue<0.05
```


```{r}
#Creating a function called CorrectMultP() that takes a vector of p-values as input and outputs columns for the p-value, and a logical column each for whether it is significant by Bonferroni, BenHotch, or qvalue.
CorrectMultP <- function(p.vect){
return(data.frame(
Bonf = Bonf(p.vect),
BenHotch = BenHotch(p.vect),
Qval = qvalue(p.vect)$qvalue<0.05))
}

out <- CorrectMultP(p.list)
head(out)
```

```{r}
### Calculate the number of significant tests
sum(out$Bonf.Is.Sig)

sum(out$BenHotch.Is.Sig)

sum(out$Qval)
```



#Checkpoint 3: Which test is the most conservative? least conservative? The true dataset only had 100 loci under selection.
#Bonferroni is the most conservative with the fewest significant result and q-value is the least conservative with most significant results.